---
title: "GDD"
author: "Nick Triozzi"
date: "6/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(riem)
library(stringr)
library(adehabitatHR)
library(rgdal)
library(gstat)
library(sp)
library(plyr)
library(dplyr)
library(pollen)
library(mapview)
library(spatstat)  # Used for the dirichlet tessellation function
library(maptools)  # Used for conversion from SPDF to ppp
library(raster)    # Used to clip out thiessen polygons

```

## Ravni Kotari Weather

This chunk retrieves weather data from riem. Here I create a table giving monthly max and min temps

```{r monthly temp extremes}
#-- retrieve weather data
library(riem)
riem_networks()
riem_stations("HR__ASOS")
# create spdf from stations, plot
hr.stations = riem_stations("HR__ASOS")
coords = data.frame(x = hr.stations$lon, y=hr.stations$lat) # create data frame of coordinates
#-- set crs for long/lat - wgs84
crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"
#-- create spdf
hrs.spdf = SpatialPointsDataFrame(coords=coords, data =hr.stations, proj4string = CRS(crs))
mapview(hrs.spdf)

# Zadar statio is LDZD; query this data
Zadar=riem_measures("LDZD", date_start = "2014-01-01", date_end = "2020-05-30")
head(Zadar)

# create date column
Zadar$date= as.Date(as.character(substr(Zadar$valid, 1, 10)), foformat = "%Y%m%d")
Zadar$year = as.numeric(format(Zadar$date, format = "%Y"))
Zadar$month = as.numeric(format(Zadar$date, format = "%m"))
Zadar$day = as.numeric(format(Zadar$date, format = "%d"))

# convert temperature to celcius
#-- convert F to Celcius
Zadar$tmpC = ((Zadar$tmpf-32)*5)/9

# get daily min and max temps
dailyTemps = Zadar %>% group_by(year,month,day) %>% summarize(minTemp=min(tmpC), maxTemp=max(tmpC))
dailyTemps

# get monthly average min and max temp for years 
monthlyTempAvg = dailyTemps %>% group_by(month) %>% summarize(meanMinTemp= mean(minTemp, na.rm=T), meanMaxTemp= mean(maxTemp, na.rm=T))
t(monthlyTempAvg)
# write.csv(t(monthlyTempAvg), file = "Zadar_monthly_Temp_20142020.csv")

```

## Create GDD raster image

Use weather station data available for RV to calculate growing degree days. Data was downloaded from 


```{r additional stations}
df = read.table("./data/weather_1952_2020.txt", stringsAsFactors = FALSE, header=T, na.strings=c('99999','99999.9','999.9','99.9'), sep = ",")
str(df)
dim(df)
colnames(df)
df = df[,c(1,3,18,19, 20)] # drop columns except for min/max temp, station ID, and date

#-- remove asterisk from temperatures and force numeric
df$TMAX = as.numeric(str_replace(df$MAX, '\\*', ''))
df$TMIN = as.numeric(str_replace(df$MIN, '\\*', ''))
df$PRCP = as.numeric(str_replace(df$PRCP, '\\*', ''))
head(df)

#-- set 9999.90 to NA
df[df$TMAX==9999.90,]$TMAX = NA
df[df$TMIN==9999.90,]$TMIN = NA

#-- drop columns
df = df[,c(1,2,5,6,7)]
colnames(df) = c("STATION_ID", "YEARMODA",  "PRCP", "TMAX","TMIN")

#-- Reformat DATE and create Year Month Day columns from NewDate column
df$date = as.Date(as.character(df$YEARMODA), format("%Y%m%d"))
df$year = as.numeric(format(df$date, format = "%Y"))
df$month = as.numeric(format(df$date, format = "%m"))
df$day = as.numeric(format(df$date, format = "%d"))
head(df)

#-- convert F to Celcius
df$TMAXC = ((df$TMAX-32)*5)/9
df$TMINC = ((df$TMIN-32)*5)/9


#-- join stations and weather data
#-- read in stations file
s = read.csv("./data/stations.csv", header = T)
str(s)
head(s)
#-- reformat station ID to match column in daily weather data
s$STATION_ID = (s$STATION_ID-99999)/100000
HR.weather = inner_join(s,df)
head(HR.weather)


```

## Subset stations in RV
Use mapview to cull stations
```{r compute GDD}
#-- group data by station and compute average daily min and max temps
HR.temps = HR.weather %>% group_by(STATION_ID, month, day) %>% summarize(meanTmax = mean(TMAXC, na.rm=T, trim=0), meanTmin=mean(TMINC, na.rm=T, trim=0))

#-- compute GDD function
source("./R/computeGDD.R")
HR.temps$GDD=compute.GDD(tmin = HR.temps$meanTmin,tmax = HR.temps$meanTmax)

#-- set NaN to 0
HR.temps[is.na(HR.temps$GDD),]=0
head(HR.temps)

#-- sum GDD per station
HR.GDD=HR.temps %>% group_by(STATION_ID) %>% summarize(GDD=sum(GDD))
head(HR.GDD)
HR.GDD=HR.GDD[-1,] # drop first empty row

#-- join to station locations
HR.GDD = inner_join(HR.GDD, s)

#-- create spatial points data frame
coords=data.frame(x = HR.GDD$LONGITUDE, y = HR.GDD$LATITUDE) # create data frame of coordinates
crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"
stations = SpatialPointsDataFrame(coords=coords, data = HR.GDD, proj4string = CRS(crs))

#-- view map
m1=mapview(stations)
m1

```

## Interpolate GDD values
First we reproject station locations to LAEA coord system

```{r interpolation}
hr = readOGR(dsn="./vectors", layer = "HRV_adm0") 
proj4string(hr)
#-- reproject to match other data - set crs to proj4string
crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
hr=spTransform(hr, CRSobj = crs)
stations = spTransform(stations, CRSobj = crs)
P=stations
W=hr

# Replace point boundary extent with that of Croatia
P@bbox = W@bbox


P.ppp <- as.ppp(P)

#-- Create an empty grid where n is the total number of cells
einkorn=raster("./rasters/einkorn_suitability.tif")
grd              = as.data.frame(spsample(P, "regular", n=ncell(einkorn), cellsize=500))
names(grd)       = c("X", "Y")
coordinates(grd) = c("X", "Y")
gridded(grd)     = TRUE  # Create SpatialPixel object
fullgrid(grd)    = TRUE  # Create SpatialGrid object
#-- Add P's projection information to the empty grid
proj4string(grd) = proj4string(P)

#-- Interpolate the grid cells using a power value of 2 (idp=2.0)
P.idw = gstat::idw(GDD ~ 1, P, newdata=grd, idp=2.0)

#-- Convert to raster object then clip to HR boundary
HR.GDD = mask(raster(P.idw), W)
plot(HR.GDD)

#-- resample and crop to study area boundary
HR.GDD.rs = resample(HR.GDD, einkorn)
plot(HR.GDD.rs)
rm(HR.GDD)
# writeRaster(HR.GDD.rs,filename =  "./rasters/HR_GDD.tif")

```

