---
title: "Gorchels_Triozzi_ESM232_final"
author: "Nick Triozzi"
date: "5/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# load functions


```

## Agricultural component

Agricultural yields have great variation in the same crop by year, location, and farming strategies This makes greating a general yield function for a specific crop challenging. To address this, we have followed a similar approach to Gulland et al. 1996 by making our agricultural modeling component adaptable. This model uses an input dataframe with at least 30 years of yield data. This can either be specific data for the farm studied or averages in the area of interest. Like Gulland et al. 1996, our function uses a normal distribution to account for the variation and gives a simulated yield for each time step based on the presented data. 

#Example wheat data
```{r crops }
source("./R/wheat_yield.R")

source("./R/barley_yield.R")
#-- function for simulating crop yield data

wheat<-data.frame(year=seq(1,50), mean_yield=rpois(50,lambda = 67))
barley<-data.frame(year=seq(1,50), mean_yield=rpois(50,lambda = 74))


example_yields=read_csv("example_yield.csv")

wheat_input = example_yields %>% 
  select(year,yield_wheat) %>% 
  rename(mean_yield=yield_wheat)

wheat_yield(yields = wheat_input)

ggplot(wheat_input)+
  geom_histogram(aes(x=mean_yield), fill="burlywood3", binwidth = 0.25)+
  theme_classic()+
  ylab("Count")+
  xlab("Yield (tonnes/ha)")+
  scale_y_continuous(expand = c(0,0))
```

# Wheat Results 
```{r}

years = seq(from=1, to=100, by=1)
yield_res= sapply(years, wheat_yield, yields = wheat_input)
# keep track of what times we ran
yield_den = data.frame(yield=yield_res, years=years)

ggplot(yield_den)+
  geom_histogram(aes(x=yield), fill="burlywood3")+
  theme_classic()+
  ylab("Count")+
  xlab("Wheat Yield (tonnes/ha)")+
  scale_y_continuous(expand = c(0,0))

```


#Barley Results 

```{r}
source("./R/barley_yield.R")

example_yields=read_csv("example_yield.csv")

barley_input = example_yields %>% 
  select(year,yield_barley) %>% 
  rename(mean_yield=yield_barley)

barley_yield(yields = barley_input)


yield_res_b= sapply(years, barley_yield, yields = barley_input)
# keep track of what times we ran
yield_den_b = data.frame(yield=yield_res_b, years=years)

ggplot(yield_den_b)+
  geom_histogram(aes(x=yield), fill="burlywood4")+
  theme_classic()+
  ylab("Count")+
  xlab("Barley Yield (tonnes/ha)")+
  scale_y_continuous(expand = c(0,0))


```


## Livestock component

We're only interested in herd growth during one time step, and input parameters, especially fertility and survivability might be modified depending on the goals of the farmer. Here, we've modified the function evolve_pop to account for additional herd reductions from slaughter. Two offtake rates are considered: 50% and 75% use rates for exploiting natural deaths for caloric gain. These figures variably affect different age categories each with its own fertility rate. Offtake for slaughter may have down-the-line consequences for herd size. 

```{r livestock, echo=FALSE}

source("./R/evolve_pop_wofftake.R")

nyears=50
nyears=nyears
# assign values for fertility for each of the age classes 
# number of daughters expected at age x from females still alive at age x
goatfert =  c(0,0.150,0.410,0.550,0.705, 0.695, 0.725, 0.515,0.515)
fertility = goatfert

# survivability - based on mortality rates female sheep (Redding Table IV-7)
goatsurv <- c(0.55, 0.90, 0.90, 0.95, 0.95, 0.90, 0.75, 0.50, 0.00)
survivability <- goatsurv

# initial population parameters
# p0 = c(33, 21, 17, 15, 13, 12, 11, 7, 4)
# sheepini = p0
# initialpop = sheepini
p0 = c(48, 23, 19, 15, 13, 12, 9, 6, 3)
goatini=p0
initialpop = goatini

# offtake rate per 100 adult does at 50% use rate of animals to die
goatoff=list(goatoff50 = c(10.8, 1.1, 0.9, 0.3, 0.3, 0.6, 1.1, 1.5 , 1.5),  # at 50%
     goatoff75 = c(16.2, 1.7, 1.4, 0.5, 0.4, 0.9, 1.6, 2.2 , 2.2)) # at 75% use rate

offtake= lapply(goatoff, function(o){o/p0}) # create offtake rates as % of initial population

# return mean offtake for each strategy
lapply(offtake,function(o){tmp=evolve_pop_wofftake(fertility, survivability, initialpop = p0, nstep = nyears, offtake=o); return(list(mean=mean(tmp[]$offtot)))})

# return mean offtake for each strategy
lapply(offtake,function(o){tmp=evolve_pop_wofftake(fertility, survivability, initialpop = p0, nstep = nyears, offtake=o); return(list(sd=sd(tmp[]$offtot)))})

tmp <- lapply(offtake,function(o){tmp=evolve_pop_wofftake(fertility, survivability, initialpop = p0, nstep = nyears, offtake=o); return(tmp)})



```



```{r plot livestock, echo=FALSE}

#-- compare population trends for both strategies 
goat50=cbind.data.frame(year=seq(1:nyears), t(tmp$goatoff50$popbyage))
goat75=cbind.data.frame(year=seq(1:nyears), t(tmp$goatoff75$popbyage))
agecats <- c("year", "0-1", "1-2", "2-3", "3-4", "4-5", "5-6", "6-7", "7-8", "8-9")
colnames(goat50)<- agecats
colnames(goat75)<- agecats
goat50_agesl= goat50 %>% gather(key="agecat", value="pop",-year)
goat75_agesl= goat75 %>% gather(key="agecat", value="pop",-year)

# plots
ggplot(goat50_agesl, aes(year, pop, fill=agecat))+geom_col()+labs(y="Population", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))

ggplot(goat75_agesl, aes(year, pop, fill=agecat))+geom_col()+labs(y="Population", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))

#-- compare offtake trends for both strategies 
goat50off=cbind.data.frame(year=seq(2:nyears), t(tmp$goatoff50$offtakes))
goat75off=cbind.data.frame(year=seq(2:nyears), t(tmp$goatoff75$offtakes))
colnames(goat50off)<- agecats 
colnames(goat75off)<- agecats
goat50_agesl_off= goat50off %>% gather(key="agecat", value="slaughtered",-year)
goat75_agesl_off= goat75off %>% gather(key="agecat", value="slaughtered",-year)

# plots

### Facet wrap woul be nice here
ggplot(goat50_agesl_off, aes(year, slaughtered, fill=agecat))+geom_col()+labs(y="Animals available for slaughter", fill="Age Group")
ggplot(goat75_agesl_off, aes(year, slaughtered, fill=agecat))+geom_col()+labs(y="Animals available for slaughter", fill="Age Group")

#-- over the 50 year period, which strategy obtains more meat?
ots <- data.frame(g50=tmp$goatoff50$offtot, g75=tmp$goatoff75$offtot )
ots <- ots[-c(1:2),] # drop first two rows
ots.g <- ots%>% gather(key="userate", value="count")


ggplot(goat50_agesl_off, aes(year, pop, fill=agecat))+
  geom_col()+
  labs(y="Animals available for slaughter", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))

ggplot(goat75_agesl_off, aes(year, pop, fill=agecat))+
  geom_col()+labs(y="Animals available for slaughter", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))


#-- A 50% use of naturally occuring herd deaths results in more animals available for offtake over a 50-year period
ggplot(ots.g, aes(userate, count)) + geom_boxplot()

```

## The Farmer's Choice Model

A 50% use of naturally occuring herd deaths results in more animals available for offtake over a 50-year period. We'll use stochastic dynamic programming (sdp) to determine if 50% use rate is the optimal strategy when paired with variable cropping strategies

Here we create lists containing the parameters for the sdp function. There are two state variables: wealth which is a measure of fitness, and age-structured goat herd population. At the start of each time step, a farmer decides whether to sow wheat (a drought sensitive crop) or barley (a drought resistant crop). The average yields are based on a poisson distribution, and is calculated as shown above. Depending on the yield, a cost to wealth is incurred reflecting seed, and is proportional to the yield. If the farmers wealth during the previous timestep is greater than the seed cost, the gains from planting either wheat or barley can be expected. The probability of obtaining any yield is based on a poisson distribution. 


```{r set up for sdp, echo=FALSE}
nyears=50
parms= list(x_crit=0, # critical mass to survive
            x_max=160, # maximal mass
            t_max=50, # number of time steps
            hsr=36,
            maxwealth=160,
            max_store=20) 

x_class=parms$x_crit:parms$x_max

#-- stock and crop parms from above
stock.parms= list(offtake=offtake, fertility=fertility, survivability=survivability, nstocks=2)
crop.parms=list(wheatY=wheat, barleyY=barley, wl=67, bl=74, ncrops=2)

# initialize outputs
times=seq(1:parms$t_max)
nclass=length(x_class)
f <- matrix(nrow=parms$t_max,ncol=nclass ,0) # optimal fitness
s <- matrix(nrow=parms$t_max,ncol=nclass ,0) # surplus
best.strategy <- matrix(nrow=parms$t_max-1,ncol=nclass-1,0) # best patch
    
# run function    
sdp_farm(f,s, times = times, parms = parms, crop.parms = crop.parms, stock.parms = stock.parms, x_class = x_class)



```

## more parameters

setting up data structures


```{r set up for sdp2, echo=FALSE}







# initialize matrices to store optimal crop and livestock strategies
best.strategy= list(crops=matrix(nrow=parms$t_max-1,ncol=nmass-1 ,0) ,
                    stock=matrix(nrow=parms$t_max-1,ncol=nmass-1 ,0))

# initialize vectors for storing current fitness 
Vs <- vector(length=parms$npatch) # current fitness
Vc <- vector(length=parms$npatch) # current fitness

f[parms$t_max,] <- 36
s[parms$t_max,] <- 20


```

Big build up

```{r deficit function, echo=FALSE}


deficit.surplus <- function(hsr, gains, maxwealth, x, t){
  # gains<- cropgains
  # maxwealth=160
  # hsr=36
  xx <- pmin(gains, maxwealth)
  d <- ifelse(xx-hsr>0, 0, xx-hsr)
  # d.index <- which.max(d)
  deficit <- d
  s <- ifelse(d==0, xx-hsr, 0)
  # s.index <- which.max(s)
  surplus <- s
  return(list(deficit=deficit, surplus=surplus))
  # f[t-1,x+1] <- surplus+x+1
}


```


```{r run the func}

##-- sdpfunction.

#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.









```

##Sensitivity analysis
goatfert =  c(0,0.150,0.410,0.550,0.705, 0.695, 0.725, 0.515,0.515)
fertility = goatfert

# survivability - based on mortality rates female sheep (Redding Table IV-7)
goatsurv <- c(0.55, 0.90, 0.90, 0.95, 0.95, 0.90, 0.75, 0.50, 0.00)
survivability <- goatsurv

```{r}
library(sensitivity)

source("./R/evolve_pop_wofftake.R")

np=100

f1=rnorm(mean = 0, sd=0.10, n=1)
f2=rnorm(mean = 0.150, sd=0.10, n=1)
f3=rnorm(mean = 0.410, sd=0.10, n=1)
f4=rnorm(mean = 0.550, sd=0.10, n=1)
f5=rnorm(mean = 0.705, sd=0.10, n=1)
f6=rnorm(mean = 0.695,sd=0.10, n=1)
f7=rnorm(mean = 0.725,sd=0.10, n=1)
f8=rnorm(mean = 0.515,sd=0.10, n=1)
f9=rnorm(mean = 0.515,sd=0.10, n=1)

fert = cbind.data.frame(f1=f1, f2=f2, f3=f3, f4=f4, f5=f5, f6=f6, f7=f7, f8=f8, f9=f9)

f1_2=rnorm(mean = 0, sd=0.10, n=1)
f2_2=rnorm(mean = 0.150, sd=0.10, n=1)
f3_2=rnorm(mean = 0.410, sd=0.10, n=1)
f4_2=rnorm(mean = 0.550, sd=0.10, n=1)
f5_2=rnorm(mean = 0.705, sd=0.10, n=1)
f6_2=rnorm(mean = 0.695,sd=0.10, n=1)
f7_2=rnorm(mean = 0.725,sd=0.10, n=1)
f8_2=rnorm(mean = 0.515,sd=0.10, n=1)
f9_2=rnorm(mean = 0.515,sd=0.10, n=1)

fert_2 = cbind.data.frame(f1=f1_2, f2=f2_2, f3=f3_2, f4=f4_2, f5=f5_2, f6=f6_2, f7=f7_2, f8=f8_2, f9=f9_2)


# create our sobel object and get sets ofparameters for running the model
sens_P = sobol2007(model = NULL, fert,fert_2, nboot =300)

#c(sens_P$X1$f1,sens_P$X1$f2,sens_P$X1$f3,sens_P$X1$f4,sens_P$X1$f5, sens_P$X1$f6, sens_P$X1$f7, sens_P$X1$f8,sens_P$X1$f9)

fertility = c(sens_P$X1$f1,sens_P$X1$f2,sens_P$X1$f3,sens_P$X1$f4,sens_P$X1$f5, sens_P$X1$f6, sens_P$X1$f7, sens_P$X1$f8,sens_P$X1$f9)

results_soble = lapply(offtake,function(o){tmp=evolve_pop_wofftake(fertility=c(sens_P$X1$f1,sens_P$X1$f2,sens_P$X1$f3,sens_P$X1$f4,sens_P$X1$f5, sens_P$X1$f6, sens_P$X1$f7, sens_P$X1$f8,sens_P$X1$f9), survivability, initialpop = p0, nstep = nyears, offtake=o); return(tmp)})

#sobol_res = mapply(FUN = evolve_pop_wofftake, 
#                   fertility=c(sens_P$X1$f1,sens_P$X1$f2,sens_P$X1$f3,sens_P$X1$f4,sens_P$X1$f5, sens_P$X1$f6, sens_P$X1$f7, sens_P$X1$f8,sens_P$X1$f9), 
#                   survivability=c(sens_P$X1$f1,sens_P$X1$f2,sens_P$X1$f3,sens_P$X1$f4,sens_P$X1$f5, sens_P$X1$f6, sens_P$X1$f7, sens_P$X1$f8,sens_P$X1$f9), 
 #                  initialpop = goatini,
  #                 nstep=nyears, 
   #                offtake=offtake) 

sens_sobol2 = sensitivity::tell(sens_sobol2, results_soble, res.names="mean_yield")

# first-order indices (main effect without co-variance)
sens_almond_sobol2$S

# total sensitivity index - note that this partitions the output variance - so values sum to 1
sens_almond_sobol2$T

print(sens_almond_sobol2)
plot(sens_almond_sobol2)

# making a datafram for plotting
both2 = cbind.data.frame(sens_almond_sobol2$X, yield=sens_almond_sobol2$y)

# Looking at the response to the two most important variables
ggplot(both2, aes(precip_coeff2, yield, col = precip_coeff)) +
  geom_point() +
  labs(x = "Precipitation Coefficient 2", y = "Yield Anamoly") +
  theme_classic()

ggplot(both2, aes(y = yield)) +
  geom_boxplot() +
  theme_classic() +
  scale_y_continuous()

```

##References 
Milner-Gulland, E.J., Mace, R., Scoones, I.1996."A Model of Household Deciscions in Dryland Agropastoral Systems".Agricultral Systems. 51:4, 407-430.


