---
title: "ESM 232 Final Project"
author: "Nick Triozzi and Madeline Gorchels"
date: "6/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sensitivity)
library(knitr)
# load functions


```

## Agricultural component

Agricultural yields have great variation in the same crop by year, location, and farming strategies This makes greating a general yield function for a specific crop challenging. To address this, we have followed a similar approach to Gulland et al. 1996 by making our agricultural modeling component adaptable. This model uses an input dataframe with at least 30 years of yield data. This can either be specific data for the farm studied or averages in the area of interest. Like Gulland et al. 1996, our function uses a normal distribution to account for the variation and gives a simulated yield for each time step based on the presented data. 

#Example wheat data
```{r crops, echo=FALSE, warning=FALSE, message=FALSE}
source("./R/wheat_yield.R")

source("./R/barley_yield.R")
#-- function for simulating crop yield data

wheat<-data.frame(year=seq(1,50), mean_yield=rpois(50,lambda = 67))
barley<-data.frame(year=seq(1,50), mean_yield=rpois(50,lambda = 74))


example_yields=read_csv("example_yield.csv")

wheat_input = example_yields %>% 
  select(year,yield_wheat) %>% 
  rename(mean_yield=yield_wheat)

#wheat_yield(yields = wheat_input)

ggplot(wheat_input)+
  geom_histogram(aes(x=mean_yield), fill="burlywood3", binwidth = 0.25)+
  theme_classic()+
  ylab("Count")+
  xlab("Yield (tonnes/ha)")+
  scale_y_continuous(expand = c(0,0))
```

# Wheat Results 

To test the function, the wheat function was run for a period of 100 years. This is an example distribution from the function, the individual input will vary once a part of the larger decision model. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}

years = seq(from=1, to=100, by=1)
yield_res= sapply(years, wheat_yield, yields = wheat_input)
# keep track of what times we ran
yield_den = data.frame(yield=yield_res, years=years)

ggplot(yield_den)+
  geom_histogram(aes(x=yield), fill="burlywood3")+
  theme_classic()+
  ylab("Count")+
  xlab("Wheat Yield (tonnes/ha)")+
  scale_y_continuous(expand = c(0,0))

```


#Barley Results 

To test the function, the barely function was run for a period of 100 years. This is an example distribution from the function, the individual input will vary once a part of the larger decision model. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("./R/barley_yield.R")

example_yields=read_csv("example_yield.csv")

barley_input = example_yields %>% 
  select(year,yield_barley) %>% 
  rename(mean_yield=yield_barley)

#barley_yield(yields = barley_input)


yield_res_b= sapply(years, barley_yield, yields = barley_input)
# keep track of what times we ran
yield_den_b = data.frame(yield=yield_res_b, years=years)

ggplot(yield_den_b)+
  geom_histogram(aes(x=yield), fill="burlywood4")+
  theme_classic()+
  ylab("Count")+
  xlab("Barley Yield (tonnes/ha)")+
  scale_y_continuous(expand = c(0,0))


```


## Livestock component

We're only interested in herd growth during one time step, and input parameters, especially fertility and survivability might be modified depending on the goals of the farmer. Here, we've modified the function evolve_pop to account for additional herd reductions from slaughter. Two offtake rates are considered: 50% and 75% use rates for exploiting natural deaths for caloric gain. These figures variably affect different age categories each with its own fertility rate. Offtake for slaughter may have down-the-line consequences for herd size. 

```{r livestock, echo=FALSE, warning=FALSE, message=FALSE}

# source("./R/evolve_pop_wofftakeK.R")
source("./R/evolve_pop_wofftakeK_play.R")
nyears=50
nyears=nyears
# assign values for fertility for each of the age classes 
# number of daughters expected at age x from females still alive at age x
goatfert =  c(0,0.150,0.410,0.550,0.705, 0.695, 0.725, 0.515,0.515)
fertility = goatfert

# survivability - based on mortality rates female sheep (Redding Table IV-7)
goatsurv <- c(0.55, 0.90, 0.90, 0.95, 0.95, 0.90, 0.75, 0.50, 0.00)
survivability <- goatsurv

# initial population parameters
p0 = c(48, 23, 19, 15, 13, 12, 9, 6, 3)
goatini=p0
initialpop = goatini


tmp1<-evolve_pop_wofftake(fertility = fertility, survivability = survivability,initialpop = p0, nstep = nyears, offtake = 0.75, K=220, minHerd = 140) 

#str(tmp)
```

## Survivability 

The number of animals available for slaughter each  year is affected by the herd's survivability. We can simulate changes to mortality rate of the herd by adjusting the parameter "extrinsicP." If extrinsicP < 1, survivability decreases, and meat availability increases; set > 1, survivability increases, and meat availability decreases. The extrinsicP parameter can be useful for modeling the effects of perturbations to the system to see how the population rebounds. For our model, extrinsicP may inform when a 75% use rate of naturally occurring deaths may fail to meet subsistence requirements when fewer animals are dying naturally.

```{r plot livestock, echo=FALSE, warning=FALSE, message=FALSE}
# test two offtake rates
offtakerate= c(0.50, 0.75)
K=320
minHerd=100
extrinsicP=0.95

tmp<-lapply(offtakerate, function(o){tmp=evolve_pop_wofftake(fertility, survivability, p0, nstep = nyears, offtake = o, extrinsicP = extrinsicP, K=K, minHerd =minHerd); return(tmp)})


# total offtake per year
yrs=seq(1:(nyears-1))
s=1:length(offtakerate)
tot.annual.offtake = sapply(s, function(s){sapply(yrs, function(y){sum(tmp[[s]]$offtakes[,y])})})
# apply(tot.annual.offtake, 2, mean)

#-- compare population trends for both strategies 
goat50=cbind.data.frame(year=seq(1:nyears), t(tmp[[1]]$popbyage))
goat75=cbind.data.frame(year=seq(1:nyears), t(tmp[[2]]$popbyage))

agecats <- c("year", "0-1", "1-2", "2-3", "3-4", "4-5", "5-6", "6-7", "7-8", "8-9")
colnames(goat50)<- agecats
colnames(goat75)<- agecats

goat50_agesl= goat50 %>% gather(key="agecat", value="pop",-year)
goat75_agesl= goat75 %>% gather(key="agecat", value="pop",-year)

# plots
ggplot(goat50_agesl, aes(year, pop, fill=agecat))+geom_col()+labs(y="Population", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))+
  ggtitle(paste0("Herd size by age at 50% use rate with K=",K," and min herd=",minHerd))+ labs(subtitle = paste0("extrinsic mortality probability=",extrinsicP ))

ggplot(goat75_agesl, aes(year, pop, fill=agecat))+geom_col()+labs(y="Population", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))+
  ggtitle(paste0("Herd size by age at 75% use rate with K=",K," and min herd=",minHerd)) + labs(subtitle = paste0("extrinsic mortality probability=",extrinsicP ))
```

## Offtake trends for both strategies 

Depending on the strategy, there may be long term offtake trends. These will lead to different availabilities of meat over time. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#-- compare offtake trends for both strategies 
goat50off=cbind.data.frame(year=seq(2:nyears), t(tmp[[1]]$offtakes))
goat75off=cbind.data.frame(year=seq(2:nyears), t(tmp[[2]]$offtakes))
colnames(goat50off)<- agecats 
colnames(goat75off)<- agecats
goat50_agesl_off= goat50off %>% gather(key="agecat", value="slaughtered",-year)
goat75_agesl_off= goat75off %>% gather(key="agecat", value="slaughtered",-year)

# plots

### Facet wrap woul be nice here
ggplot(goat50_agesl_off, aes(year, slaughtered, fill=agecat))+
  geom_col()+
  labs(y="Animals available for slaughter", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))+
  ggtitle(paste0("Slaughter by age at 50% use rate with K=",K," and min herd=",minHerd))+
  labs(subtitle = paste0("extrinsic mortality probability=",extrinsicP ))


ggplot(goat75_agesl_off, aes(year, slaughtered, fill=agecat))+
  geom_col()+labs(y="Animals available for slaughter", fill="Age Group")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))+
  ggtitle(paste0("Slaughter by age at 75% use rate with K=",K," and min herd=",minHerd))+
  labs(subtitle = paste0("extrinsic mortality probability=",extrinsicP ))

```

## Over the 50 year period, which strategy obtains more meat?

First the annual total of animals consumed shows that in general, the 75 offtake strategy has more animals consumed typically in a single year. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#-- over the 50 year period, which strategy obtains more meat?
s <- c(1,2)
y <- seq(1:ncol(tmp[[1]]$offtakes))
ot.yr <- sapply(s, function(s){sapply(yrs, function(y){sum(tmp[[s]]$offtakes[,y])})})
ot.df <- as.data.frame(ot.yr)
colnames(ot.df) <- c("g50", "g75")
ot.yr <- ot.df%>% gather(key="userate", value="annual.total")

ggplot(ot.yr, aes(userate, annual.total)) + geom_boxplot() + 
  labs(title ="Total offtakes per year for each strategy", subtitle =paste0("extrinsic mortality probability=",extrinsicP,"; K=",K," and min herd=",minHerd))+
  theme_classic()+
  xlab("Use Rate")+
  ylab("Annual Total")
```

## The offtake rates used here reflect the percentage of naturally occurring deaths used for meat. 

A function to compute average number of animals slaughtered per year for the 50% and 75% strategies.

```{r sensitivity livestock, include=FALSE}
compute.offtake.metrics <- function(res){
  s=1:length(res)
  yrs=1:ncol(res[[1]]$offtakes)
  tot.annual.offtake = sapply(s, function(s){sapply(yrs, function(y){sum(tmp[[s]]$offtakes[,y])})})
  avg.annual.offtake = sapply(s, function(s){mean(tot.annual.offtake[,s])})
  return(avg.annual.offtake)
  # return(tot.annual.offtake)
}
compute.offtake.metrics(tmp)

```

```{r Livestock sensitivity analysis, include=FALSE}
source("./R/evolve_pop_wofftakeK_play.R")

# sensitivity of herd size to 

np=100
Off <- rnorm(mean = 0.5, sd=0.25, n=np)
K <- rnorm(mean = 320, sd=60, n=np)
minHerd <- rnorm(mean = 160, sd=30, n=np)
extrinsicP<-rnorm(mean = 0.85, sd=0.15, n=np)
X1 = cbind.data.frame(Off=Off, K=K, minHerd=minHerd, extrinsicP=extrinsicP)

np=100
Off <- rnorm(mean = 0.5, sd=0.25, n=np)
K <- rnorm(mean = 320, sd=60, n=np)
minHerd <- rnorm(mean = 160, sd=30, n=np)
extrinsicP<-rnorm(mean = 0.85, sd=0.15, n=np)
X2 = cbind.data.frame(Off=Off, K=K, minHerd=minHerd,  extrinsicP=extrinsicP)

# null model
sens_P_livestock = sobol2007(model = NULL, X1,X2, nboot =300)

# wrapper function to summarize average number of animals slaughtered per year
avg.offtakes <- function(res){
  alist <- res[3,]
  annavgs=lapply(alist, function(a){ann.totals=mean(apply(a,2,FUN = sum))})
  return(annavgs)
}

res = unlist(avg.offtakes(mapply(FUN = evolve_pop_wofftake, offtake=sens_P_livestock$X$Off, K=sens_P_livestock$X$K, minHerd=sens_P_livestock$X$minHerd, extrinsicP=sens_P_livestock$X$extrinsicP, MoreArgs = list(fertility, survivability, p0, nstep=nyears)))) #; return(compute.offtake.metrics(tmp))})

sens_P_livestock = sensitivity::tell(sens_P_livestock, unlist(res), res.names="avg.ann.offtake")

#sens_P_livestock$S
#sens_P_livestock$T
```

## Livestock Sensitivity Analysis

Sensitivity analysis shows our livestock model is about equally sensitive to all input parameters. All evaluated parameters are within the margin of error with eachother. 

However, graphing the offtake rate against average annual animals consumed shows there could be a stronger relationship than the minimum heard size. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#print(sens_P_livestock)
plot(sens_P_livestock, ylim = c(-1,2))

# making a dataframe for plotting
both = cbind.data.frame(sens_P_livestock$X, results=sens_P_livestock$y)

# look at response of conductance to the two most important variables
ggplot(both, aes(extrinsicP,results, col=Off))+geom_point()+labs(y="Average Animals Slaughtered Annually", x="Probability of Extrinsic Mortality") + ggtitle("Results of sensitivity analysis of results of livestock submodel")+
  theme_classic()

# Looking at the response to the two most important variables
ggplot(both, aes(Off, results, col = extrinsicP)) +
  geom_point() +
  labs(title ="Results of sensitivity analysis of results of livestock submodel",  x = "Offtake Rate", y ="Average Animals Slaughtered Annually") +
  theme_classic()

```

## The Farmer's Choice Model

A 75% use of naturally occuring herd deaths results in more animals available for offtake over a 50-year period. We'll use stochastic dynamic programming (sdp) to determine if 50% use rate is the optimal strategy when paired with variable cropping strategies

Here we create lists containing the parameters for the sdp function. There are two state variables: wealth which is a measure of fitness, and age-structured goat herd population. At the start of each time step, a farmer decides whether to sow wheat (a drought sensitive crop) or barley (a drought resistant crop). The average yields are based on a poisson distribution, and is calculated as shown above. Depending on the yield, a cost to wealth is incurred reflecting seed, and is proportional to the yield. If the farmers wealth during the previous timestep is greater than the seed cost, the gains from planting either wheat or barley can be expected. The probability of obtaining any yield is based on a poisson distribution. 


```{r sdp model, echo=FALSE, warning=FALSE, message=FALSE}
nyears=50

parms= list(x_crit=0, # critical mass to survive
            x_max=75, # maximal mass
            t_max=50,
            maxwealth=75) # number of time steps) 


#-- stock and crop parms from above
stock.parms= list(offtake=offtakerate, fertility=fertility, survivability=survivability, nstocks=2, initialpop=p0)
crop.parms=list(wheatY=wheat, barleyY=barley, wl=67, bl=74, ncrops=2)

 source("./R/sdp_farm123.R")
# source("./R/sdp_farm123.R")
# run function    
sdp.out=sdp_farm(parms = parms, crop.parms = crop.parms, stock.parms = stock.parms)
```


```{r summarize sdp res, echo=FALSE, warning=FALSE, message=FALSE}
source("./R/sdp_df.R")
source("./R/compute_wealth_metrics.R")
```

```{r summarize sdp res2, echo=FALSE, warning=FALSE, message=FALSE}
sdp_df=function(sdp_output){
  fw=sdp_output[[1]]
  year= 1:sdp_output[[5]]
  avg.wealth=apply(fw,1 ,mean)
  sd.wealth=apply(fw,1 ,sd)
  wheat.yld=sdp_output[[6]]
  barley.yld=sdp_output[[7]]
  gp<-t(apply(sdp_output[[4]], c(1,2), sum))
  mean.herd=apply(gp,1,mean)
  
  df=data.frame(year=year,avg.wealth=avg.wealth, sd.wealth=sd.wealth, wheat.yld=wheat.yld, barley.yld=barley.yld, mean.herd=mean.herd)
  return(df)
}

# as a dataframe
farm.df=sdp_df(sdp.out)
#head(farm.df)

gg1<-ggplot(farm.df[-50,], aes(x=year,y=avg.wealth))+geom_line() 
gg1 + geom_point(aes(y=wheat.yld), color="red") +
  geom_point(aes(y=barley.yld), color="blue") +
  theme_classic()+
  labs(x="Year", y="Average Wealth")


compute.wealth.metrics= function(df, mean.only=FALSE){
  df=df[-1,]
  
  if(mean.only==TRUE){
    return(mean.wealth=mean(df$avg.wealth, na.rm = T))
  } else {
    return(list(mean.wealth=mean(df$avg.wealth, na.rm = T),
                variance=var(df$avg.wealth, na.rm = T),
                cor.wheat=cor(df$wheat.yld, df$avg.wealth),
                cor.barley=cor(df$barley.yld, df$avg.wealth),
                cor.herd= cor(df$mean.herd, df$avg.wealth) ))
  }
}

```

This graph shows the wheat yield in red, the barley yield in blue and the average wealth with the black line. This deciion model shows destitution can occur with only slight variation in crop yields. 

```{r, echo=FALSE}
compute.wealth.metrics(farm.df)
compute.wealth.metrics(farm.df, mean.only = T)


```


```{r, include=FALSE}
# plot
source("./R/plot_wealth.R")
plot.wealth <- function(sdp.out, farm.yrly){
  f1<-sdp.out$FarmWealth
  bestpatch <- sdp.out$OptimalStrategy
  wealthclasses=0:ncol(f1[,-1])
  timesteps = 0:nrow(bestpatch)


  w=0:ncol(sdp.out$FarmWealth[,-1])
  t=1:nrow(sdp.out$OptimalStrategy)
  par(xpd = T, mar = par()$mar + c(0,0,0,12))
  
  image(x=timesteps,y=wealthclasses[-1],z=bestpatch,ylab="Wealth",xlab="Time",zlim=c(0,5), main="Best Strategy",col=c("black",topo.colors(4))) 
  box() 
  
  contour(x=1:parms$t_max,y=wealthclasses,z=f1,add=TRUE, col = "white")
 
  # abline(h = mean(farm.yrly$avg.wealth), col="red", lwd=3, lty=2)
  legend(55, 25, fill=c("black",topo.colors(4), "red"), legend=c("destitue","wheat & 50% use rate","wheat & 75% use rate","barley & 50% use rate", "barley & 75% use rate", "mean wealth"))
   ablineclip(h = mean(farm.yrly$avg.wealth), x1 = 0,x2 = 49, lwd=3,lty = 2, col = "red")
  
}

plot.wealth(sdp.out, farm.df)

```

## Varying the parameters 

# Here we change the mean yield for wheat and barley in the mean.ylds variable from 67 and 74 to 58 and 64, respectively. We also see what happens when the household subsistence requirement is set to 25% of the default, 36. In the function, the mean yields are used as the lambda inputs in a poisson distribution. These values are used to generate the probabilities associated with each year's yield.

The output suggests that wheat yields were sufficiently high in the first three time periods, and a 50% goat harvest rate allowed less wealthy households to survive. The average wealth for the run is higher than previously. The second plot shows that average wealth decreased according to drops in barley and wheat harvest between years 10 and 20. 

```{r changing params, echo=FALSE, warning=FALSE, message=FALSE}
crop.parms=list(wheatY=wheat, barleyY=barley, ncrops=2)
sdp.out=sdp_farm(parms = parms, crop.parms = crop.parms, mean.ylds = c(58,64), stock.parms = stock.parms, hsr = 9)


# show optimal strategy
farm.df=sdp_df(sdp.out)
plot.wealth(sdp.out, farm.df)

# plot yield effects on wealth
gg1<-ggplot(farm.df[-50,], aes(x=year,y=avg.wealth))+geom_line() 
gg1 + geom_point(aes(y=wheat.yld), color="red") +
  geom_point(aes(y=barley.yld), color="blue") +
  theme_classic()+
  labs(x="Year", y="Average Wealth")


```

## Sensitivity Analysis for SDP 

```{r sensitivity for sdp, echo=FALSE}
# Set up for sobol sensitivity analysis
np=100
hsr = rnorm(mean=36,sd=12, n=np)
max_store = runif(min=12,max=30, n=np)
PrS = rnorm(mean=0.85, sd=0.30, n=np)
PrGood = rnorm(mean=0.8, sd=0.25, n=np)
D.good = rnorm(mean=0.75, sd=0.15, n=np)
D.bad = rnorm(mean=0.85, sd=0.35, n=np)
X1 = cbind.data.frame(hsr, max_store, PrS, PrGood, D.good, D.bad)

# repeat sampling
np=100
hsr =  rnorm(mean=36,sd=12, n=np)
max_store = runif(min=12,max=30, n=np)
PrS = rnorm(mean=0.85, sd=0.30, n=np)
PrGood = rnorm(mean=0.8, sd=0.25, n=np)
D.good = rnorm(mean=0.75, sd=0.15, n=np)
D.bad = rnorm(mean=0.85, sd=0.35, n=np)
X2 = cbind.data.frame(hsr, max_store, PrS, PrGood, D.good, D.bad)

# sobol 2002 method used here
sens_sdp_sobel = sobol2002(model = NULL, X1, X2, nboot = 100)

# run model for all parameter sets
sdp_res = mapply(FUN=sdp_farm,hsr=sens_sdp_sobel$X$hsr,max_store=sens_sdp_sobel$X$max_store,PrS=sens_sdp_sobel$X$PrS, PrGood=sens_sdp_sobel$X$PrGood,D.good=sens_sdp_sobel$X$D.good, D.bad=sens_sdp_sobel$X$D.bad, MoreArgs=list(parms=parms, crop.parms=crop.parms, stock.parms=stock.parms))

# wrapper function for summarizing sdp results
wrapper<- function(res){
  df <- sdp_df(res)
  results<-compute.wealth.metrics(df, mean.only = T)
  return(results)
}

# apply wrapper to results
sdp_results=sapply(1:dim(sdp_res)[2], function(r){wrapper(sdp_res[,r])})

# feed results back to sensitivity model
sens_sdp_sobel = tell(sens_sdp_sobel,sdp_results, res.names="avg.w")

# first-order indices (main effect without co-variance)
sens_sdp_sobel$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
sens_sdp_sobel$T

print(sens_sdp_sobel)

# compare effects of params
plot(sens_sdp_sobel, ylim=c(-1,2))

# make a data frame for plotting
both = cbind.data.frame(sens_sdp_sobel$X, av.w=sens_sdp_sobel$y)

# look at response of conductance to the two most important variables
ggplot(both, aes(hsr,av.w, col=PrS))+geom_point()+labs(y="Average Wealth", x="HSR")


```

##Conclusions

Our findings match real world subsistance farming patterns, showing much variability due to the variation in yields. However, there were some useful takeaways. First, while offtaking animals at the 75% rate can generate more meat in a given year, offtaking animals at 50% leads to overall more meat on a longer time scale, but relying on this strategy can fail when the mortality of the herd deviates from expectation. Sensitivity analysis on the livestock function shows no significant differences between parameters, but visualizaion indicates there could be a stronger relationship between offtake rates and average annual meat consumption. 



##References 
Milner-Gulland, E.J., Mace, R., Scoones, I.1996."A Model of Household Deciscions in Dryland Agropastoral Systems".Agricultral Systems. 51:4, 407-430.


